---
title: "NDH802 - Assignment 1"
author: "Group no. 10 "
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

------------------------------------------------------------------------

-   The assignment includes 2 questions, 6 points each.
-   Submit your assignment via Canvas before 15:00 CET, April 6, 2022.
-   Your submission should be an RMarkdown file with your solutions in words and/or R code. If you handwrite parts of your assignment, insert it as an image near the corresponding question(s). Name the files NDH802_Assignment1_GroupNumber.
-   You should work in groups and contribute equally.
-   You can copy my code, but make sure you understand it.
-   You should not have the exact solutions and/or results with other groups.
-   Results without code/justifications will not be graded.

------------------------------------------------------------------------

### Set things up

Set your working directory and fill in your group number. For example, if you are group 3, make it `our_group <- 3`. If you don't fill in your group number or fill in the wrong number, your assignment will **not** be graded.

```{r}
#setwd("")
our_group <- 10
```

Run this code chunk to load data into your R Environment. The command will randomly select 1,000,000 rows of data from the original data set. Hereby each and every group should have a unique `df`. Accordingly, your results should be different from other groups' and you should not be comparing them.

```{r, echo=FALSE}
inference_dataset <- read.csv("https://cda.hhs.se/NDH802data2022.csv")
set.seed(our_group); df <- inference_dataset[sample(1:nrow(inference_dataset),
                           size = 1000000,
                           replace = FALSE), -1]; rm(inference_dataset)
```

#### Data description

+-------------------+------------------------------------------------------------------------------------------------------------+
| Variable          | Description                                                                                                |
+===================+============================================================================================================+
| cust.id           | Unique customer id                                                                                         |
+-------------------+------------------------------------------------------------------------------------------------------------+
| age               | Customer age in year 2021                                                                                  |
+-------------------+------------------------------------------------------------------------------------------------------------+
| email             | If there is an email of the customer in the system                                                         |
+-------------------+------------------------------------------------------------------------------------------------------------+
| member.since      | Year from which the customer become a member. They can only register at the physical stores.               |
+-------------------+------------------------------------------------------------------------------------------------------------+
| distance.to.store | Distance (in km) from customer's address to the physical store they register their membership              |
+-------------------+------------------------------------------------------------------------------------------------------------+
| store.trans       | Total number of offline transactions the customer made in year 2021                                        |
+-------------------+------------------------------------------------------------------------------------------------------------+
| store.spend       | Total amount (in SEK) the customer spend from offline transaction in year 2021                             |
+-------------------+------------------------------------------------------------------------------------------------------------+
| online.visits     | Total number of time the customer visit (does not necessarily mean purchase) the online store in year 2021 |
+-------------------+------------------------------------------------------------------------------------------------------------+
| online.trans      | Total number of online transactions the customer made in year 2021                                         |
+-------------------+------------------------------------------------------------------------------------------------------------+
| online.spend      | Total amount (in SEK) the customer spend from online transactions in year 2021                             |
+-------------------+------------------------------------------------------------------------------------------------------------+
| points            | Total loyalty points the customer accumulates since they become a member deducted by points they have used |
+-------------------+------------------------------------------------------------------------------------------------------------+
| main.format       | | The format in which the customer made the most transactions in year 2021                                 |
+-------------------+------------------------------------------------------------------------------------------------------------+

\newpage

### Question 1. Mean and variance

(a) Compute the `total.spend` of the customer (that includes `store.spend` and `online.spend`). Plot the histogram of `total.spend`. Imagine you will present this to your manager. Make it readable and self-explanatory (e.g., add the title for the chart and labels for the axes where needed). (1p) How do you explain the peak of the histogram, in general and in this context? (1p)

    **Answer:** The peak area of the histogram constitutes the spend of which the majority of customers are putting on both online and store items.

    In general, the peaks show where the majority of responses or values are located at. For example, if the most students scored 10 on a test this will be the highest peak.

    ```{r}
    hist(
      df[,"total.spend"],
      main = "Total Spend",
      xlab = "Sum of Store and Online Spend",
      ylab = "No. of Customers",
      col = "lightblue3"
    )
    ```

(b) Make a box plot for `total.spend` for 3 groups, customers whose main format is supermarket, convenience store and online. Refer to the code provided and modify it (1p). From this figure, would you conclude that online is the format that contributes the most to the `total.spend`? Why/why not? (1p)

    **Answer:** Upon analyzing the boxplot we cannot draw the conclusion that the online format contributes the most to online spend. However, we can see the variability regarding spend in the diagram below. The boxplot does not show the quantity of these transactions, rather it shows what values, in term of spend, is usually executed by the customer. Even though it might seem that one format has an higher median spend, it does not account for the number of transactions.

    Moreover, sometimes we would like to eliminate outliers but in the diagram below all entries are presented.

```{r,fig.height=6, fig.width=6}
boxplot(
  df[,"total.spend"] ~ main.format,
  data = df,
  ylim = c(10000, 350000),
  main = "Total Spend",
  xlab = "Format",
  ylab = "Spend",
  col = "lightblue3"
)
```

(c) Compute the mean and variance of `distance.to.store` of the customers whose `main.format` is supermarket and of the customers whose `main.format` is convenient store (1p). Comment on the difference between the means of the two groups; and the difference between the variances of the two groups (1p).

    **Answer:** When comparing the mean distance to store for supermarket and convenience store customers we can identify that the mean distance to store is 4,78 distance units for supermarket customer and 1,35 distance units for convenience store customers. Thus customers visiting the supermarket are driving longer distances to access greater assortment.

    We identifed that the supermarket customers had a variability of 33,65 distance units whilst convenience store customers had a variability of 0,4. We can assume that the reasons correspond with the statement above.

    ```{r}
    #Supermarket
    mean(df[df$main.format == "Supermarket", "distance.to.store"],
         na.rm = TRUE)

    var(df[df$main.format == "Supermarket", "distance.to.store"],
        na.rm = TRUE)

    #Convinience Store
    mean(df[df$main.format == "Convenience store",
                              "distance.to.store"],
                                na.rm = TRUE)

    var(df[df$main.format == "Convenience store",
                              "distance.to.store"],
                               na.rm = TRUE)
    ```

### Question 2. Probability theory

Consider the following events:

> (A) Made at least one offline transaction
>
> (B) Made at least one online transaction

(a) Compute $P(A)$ and $P(B)$. (1p)

    ```{r}
    #Assigning the probability of A to P_A:
    P_A <- sum(df$store.trans != 0) / sum(nrow(df))
    print(P_A)

    #Assigning the probability of B to P_B:
    P_B <- sum(df$online.trans != 0) / sum(nrow(df))
    print(P_B)
    ```

(b) What is the complement of $B$? Formally, define event $\bar{B}$ and compute $P(\bar{B})$. (1p)

    **Answer:** The complement of event B defines all outcomes that does not adhere to event B. In essence, we are looking for the outcomes/probability where the customers have not carried out an online transaction.

    This is calculated using the formula: $P(\bar{B})$ = $1 - P(B)$

    ```{r}
    #Uswing the above assigned variable of P_B and following the formula:
    P_Bbar <- 1 - P_B

    print(P_Bbar)
    ```

(c) Compute $P(B \cap A)$ and $P(B \mid A)$. (1p)

    ```{r}
    #Computing $P(B \cap A)$ requires us to find the intersect of where both P_A and P_B are > 0:
    P_A_union_B <- prop.table(table(df$store.trans != 0 &
                      df$online.trans != 0))

    #We are only interested in the true value as that shows where P_A and P_B intersect as displayed in a Venn diagram.
    print(P_A_union_B)

    #Computing $P(B \mid A)$ requires us to take the intersection found in above and divide it by P_A:
    P_A_union_B / P_A
    print(P_A_union_B / P_A)
    #We will also look at the true values here
    ```

(d) Are $A$ and $B$ independent events? Why/why not? (1p)

    **Answer:** If A and B are independent events then $P (A \cap B)$ = $P (A)$ \* $P (B)$. We know that $P (A \cap B)$ = 0,960125 and that $P (A)$ = 1 and that $P (B)$ = 0,960125.

    Thus, multiplying $P (A)$ with $P (B)$ should give 0,960125. Utilizing the numbers above we can see that the formula holds true. This means that A and B are statistically independent.

(e) Compare and explain (with formula) the similarities/differences among $P(B)$, $P(B \cap A)$ and $P(B \mid A)$ (2p).

    **Answer:** $P (B)$ is the probability that event B will happen in the sample space. We calculate $P (B)$ by dividing the number of favorable outcomes, regarding B by the entire sample space: $P (B) = \frac{N_B}{N}$. The probability will range from 0 to 1 where 1 represents 100% probability.

    $P (B \cap A)$ showcases when an outcome is placed both within the probability of event A and B. In essence, we are looking for convergence when eg. an individual from Group A also conforms to Group B. When regarding a Venn diagram, the intersection is placed where P(A) intersects with P(B).

    $P (B \mid A)$ shows the probability of B given that the outcome of A already occurred. We calculate this using the formula: $P(B \mid A)$ = $P (A \cap B)$ / $P (A)$ .

\vfill

*Have fun and good luck!*\
*Huong and Emelie*

## Fantastic!

I love it that you comment a long the way. It is enjoyable reading your thought process.

Q1ab. Beautifully reasoned. 4p

Q1c. Correct calculation and comment on the mean. However I think the comment on the variance is not well motivated. 1.5p

Q2abc. Nicely done 3p

Q2c. You gave me a hard time reading this answer because you write intersect in the comment and union in the code... 1p

Q2e. Everything you wrote is correct however has not explain why $P(B) = P(B \cap A) = P(B \mid A)$ in this case. 0.5
