---
title: "NDH802 - Assignment 1"
author: "Group no. 2 "
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

------------------------------------------------------------------------

-   The assignment includes 2 questions, 6 points each.
-   Submit your assignment via Canvas before 15:00 CET, April 6, 2022.
-   Your submission should be an RMarkdown file with your solutions in words and/or R code. If you handwrite parts of your assignment, insert it as an image near the corresponding question(s). Name the files NDH802_Assignment1_GroupNumber.
-   You should work in groups and contribute equally.
-   You can copy my code, but make sure you understand it.
-   You should not have the exact solutions and/or results with other groups.
-   Results without code/justifications will not be graded.

------------------------------------------------------------------------

### Set things up

Set your working directory and fill in your group number. For example, if you are group 3, make it `our_group <- 3`. If you don't fill in your group number or fill in the wrong number, your assignment will **not** be graded.

```{r}
#setwd("")
our_group <- 2
```

Run this code chunk to load data into your R Environment. The command will randomly select 1,000,000 rows of data from the original data set. Hereby each and every group should have a unique `df`. Accordingly, your results should be different from other groups' and you should not be comparing them.

```{r, echo=FALSE}
inference_dataset <- read.csv("https://cda.hhs.se/NDH802data2022.csv")
set.seed(our_group); df <- inference_dataset[sample(1:nrow(inference_dataset),
                           size = 1000000,
                           replace = FALSE), -1]; rm(inference_dataset)
```

#### Data description

+-------------------+------------------------------------------------------------------------------------------------------------+
| Variable          | Description                                                                                                |
+===================+============================================================================================================+
| cust.id           | Unique customer id                                                                                         |
+-------------------+------------------------------------------------------------------------------------------------------------+
| age               | Customer age in year 2021                                                                                  |
+-------------------+------------------------------------------------------------------------------------------------------------+
| email             | If there is an email of the customer in the system                                                         |
+-------------------+------------------------------------------------------------------------------------------------------------+
| member.since      | Year from which the customer become a member. They can only register at the physical stores.               |
+-------------------+------------------------------------------------------------------------------------------------------------+
| distance.to.store | Distance (in km) from customer's address to the physical store they register their membership              |
+-------------------+------------------------------------------------------------------------------------------------------------+
| store.trans       | Total number of offline transactions the customer made in year 2021                                        |
+-------------------+------------------------------------------------------------------------------------------------------------+
| store.spend       | Total amount (in SEK) the customer spend from offline transaction in year 2021                             |
+-------------------+------------------------------------------------------------------------------------------------------------+
| online.visits     | Total number of time the customer visit (does not necessarily mean purchase) the online store in year 2021 |
+-------------------+------------------------------------------------------------------------------------------------------------+
| online.trans      | Total number of online transactions the customer made in year 2021                                         |
+-------------------+------------------------------------------------------------------------------------------------------------+
| online.spend      | Total amount (in SEK) the customer spend from online transactions in year 2021                             |
+-------------------+------------------------------------------------------------------------------------------------------------+
| points            | Total loyalty points the customer accumulates since they become a member deducted by points they have used |
+-------------------+------------------------------------------------------------------------------------------------------------+
| main.format       | | The format in which the customer made the most transactions in year 2021                                 |
+-------------------+------------------------------------------------------------------------------------------------------------+

\newpage

### Question 1. Mean and variance

(a) Compute the `total.spend` of the customer (that includes `store.spend` and `online.spend`). Plot the histogram of `total.spend`. Imagine you will present this to your manager. Make it readable and self-explanatory (e.g., add the title for the chart and labels for the axes where needed). (1p) How do you explain the peak of the histogram, in general and in this context? (1p)
(b) Make a box plot for `total.spend` for 3 groups, customers whose main format is supermarket, convenience store and online. Refer to the code provided and modify it (1p). From this figure, would you conclude that online is the format that contributes the most to the `total.spend`? Why/why not? (1p)

```{r}
# Question a
# The peak shows how much most customers spend. Most of the customers have a low total spend of around 50.000. In general the distribution in a histogram is scewed right. 
df$total.spend=df$store.spend+df$online.spend

hist(df[, "total.spend"],
     xlim=c(0,300000),
     breaks = 100, main= "Expected spend",
     xlab = "Total spend",
     col = "pink")

```

```{r,fig.height=6, fig.width=6}
# question b
# Yes, firstly it has the highest 'minimum total spend. Secondly online stores has the highest median spend and thirdly the 25% data set above the median has the highest expenditure in the  online format.

boxplot(
total.spend ~ main.format,
data = df,
ylim = c(1, 200000), 
xlab = "Format",
ylab = "Total Spend",
col = "yellow")

```

(c) Compute the mean and variance of `distance.to.store` of the customers whose `main.format` is supermarket and of the customers whose `main.format` is convenient store (1p). Comment on the difference between the means of the two groups; and the difference between the variances of the two groups (1p).

**Mean for supermarket and convenience store**

```{r}
mean(df[df$main.format == "Supermarket", "distance.to.store"], na.rm=TRUE)
```

```{r}
mean(df[df$main.format == "Convenience store", "distance.to.store"], na.rm=TRUE)
```

**Variance for supermarket and convenience store**

```{r}
var(df[df$main.format == "Supermarket", "distance.to.store"], na.rm=TRUE)
```

```{r}
var(df[df$main.format == "Convenience store", "distance.to.store"], na.rm=TRUE)
```

```{r}
#Answer question c
#It makes sense that the supermarket has a longer mean distance than the convince stores since they are usually located further away. The purpose of a convince store is to be close. It's also reasonable that the supermarket has a higher variance than the convince stores, since people usually just want to go there when they live nearby.
```

### 

### Question 2. Probability theory

Consider the following events:

> (A) Made at least one offline transaction
>
> (B) Made at least one online transaction

(a) Compute $P(A)$ and $P(B)$. (1p)

    ```{r}
    #P(A) calculated below:
    prop.table(table(df$store.trans != 0))
    ```

    ```{r}
    #Shows that every customer in the register has carried out an offline transaction thus making the P(B) calculated below:

    prop.table(table(df$online.trans != 0))
    ```

(b) What is the complement of $B$? Formally, define event $\bar{B}$ and compute $P(\bar{B})$. (1p)

    ```{r}
    # 1 - probability = complement
    #The first method gives:
    1 - 0.959939

    ```

(c) Compute $P(B \cap A)$ and $P(B \mid A)$. (1p)

    ```{r}
    #Computing P(B intersect A):
    prop.table(table(df$store.trans !=0 & df$online.trans !=0))
    ```

    ```{r}
    #Computing p(B given A)
    #Calculating the intersect as above:
    prop.table(table(df$store.trans !=0 & df$online.trans !=0))

    #Calculting the probability of A:
    prop.table(table(df$store.trans != 0))
    ```

(d) **Are A and B independent events? Why/why not? (1p)** **Answer:** Events are independent if p(A∩B) = p(A) \* P(B).

```{=html}
<!-- -->
```
    P(A) = 1
    P(B) = 0,959939

    Probability of p(A∩B) = 0,959939, thus they are independent events

(e) **Compare and explain (with formula) the similarities/differences among P(B), P(B ∩ A) and P(B \| A) (2p)**

    **Answer:** P(B) is the probability that event b occurs in a sample data set. P(B) = number of successful trials (b occurs) / total data set.

    P(B ∩ A) is when the both event A and B occurs. Assuming independent events. P(B ∩ A) = P(B)\*P(A)

    P(B \| A) = P(**B ∩ A)/P(A),** which is the probability of B occurring given that A has occurred.

    A similarity is that in conditional probability you also incorporate P(B ∩ A) in the formula for independent events.

\vfill

*Have fun and good luck!*\
*Huong and Emelie*

## Good job!

Some minor comments:

Q1a: You should have said the peak is the mode of total spend and explain the meaning of the mode in general and in this data. I do not suggest you give the histogram the title "Expected spend", because "Expected" usually refer to the mean of the distribution. Here I still give you full score but I would like you to understand these key things. 2p

Q1b. You get 1p for the plot and no point for the second sub-question. Let's discuss this in the course because it is crucial.

Q1c. You get 1p for the calculation and 0.5 for the argument relating the means.

Q2a. Correct. 1p

Q2b. You have not defined the event $\bar{B}$. 1 - p(B) is the property, not definition. 0.5p

Q2c. You have not compute $P(B \mid A)$. 0.75p

Q2d. Correct. 1p

Q2e. Your answer does not explain why they three are similar. 1p.
