---
title: "NDH802 - Assignment 1"
author: "Group no. "
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

------------------------------------------------------------------------

-   The assignment includes 2 questions, 6 points each.
-   Submit your assignment via Canvas before 10:00 CET, April 12, 2021.
-   Your submission should be an RMarkdown file with your solutions in words and/or R code. If you handwrite parts of your assignment, insert it as an image near the corresponding question(s). Name the files NDH802_Assignment1_GroupNumber.
-   You should work in groups and contribute equally.
-   You can copy my code, but make sure you understand it.
-   You should not have the exact solutions and/or results with other groups.
-   Results without code/justifications will not be graded.

------------------------------------------------------------------------

### Set things up

Set your working directory

```{r}
#setwd("")
```

Run this code chunk to load data into your R Environment. The command will randomly select 1,000,000 rows of data from the original data set. [Fill in your group number within \`set.seed()\`.]{.ul} For example, if you are group 3, make it `set.seed(3)`. Hereby each and every group should have a unique `df`. Accordingly, your results should be different from other groups' and you should not be comparing them.\
*Note*. If you make the wrong seed, your assignment will **not** be graded.

```{r}
inference_dataset <- read.csv("https://cda.hhs.se/inference_dataset.csv")
set.seed(1); df <- inference_dataset[sample(1:nrow(inference_dataset),
                           size = 1000000, replace = FALSE), -1]
rm(inference_dataset)
summary(inference_dataset)
```

Please refer to Canvas, Hand-in 1 for more details about the data set.

\newpage

### Question 1. Mean and variance

(a) Plot the histogram of customer's `value`. Imagine you will present this to your boss at work. Make it readable and self-explanatory (e.g., add the title for the chart and labels for the axes where needed). (1p)

    ```{r}
    hist()
    ```

(b) Make a box plot for `visits` for 2 groups, loyal and not loyal customers. Refer to the code provided and modify it (1p). Imagine you are the customer relationship manager. What would you say about this figure? (1p)

    This plot means

```{r,fig.height=6, fig.width=6}
# boxplot(
#   your_variable_of_interest ~ the_group,
#   data = your_df,
#   ylim = c(-10, 100) # adjust the ylim that better illustrates your data
# )
```

(c) Compute the mean and variance of `value`, `deals`, `points`. Be careful, variance is **different from** standard deviation. (1p)
(d) Compute the mean and variance of `value`, `deals`, `points` of the *loyal* customers. (1p)
(e) Compute the mean and variance of `value`, `deals`, `points` of the *loyal* customers who *made at least one offline purchase*. (1p)

### Question 2. Probability theory

(a) How many loyal customers and not loyal customer do you have in your `df`? Formally, compute $N_{loyal}$ and $N_{\overline{loyal}}$ . (1p)

| For Q2b-f, consider the following events in your `df`:

> (E1) Being loyal
>
> (E2) Not being loyal
>
> (E3) Made at least one offline purchase
>
> (E4) Made at least one online purchase

(b) Are E1 and E2 mutually exclusive? Why/why not? (1p)
(c) Are E1 and E2 collectively exhaustive? Why/why not? (1p)
(d) Are E3 and E4 mutually exclusive? Why/why not? (1p)
(e) Are E3 and E4 collectively exhaustive? Why/why not? (1p)
(f) Are E1 and E4 mutually exclusive? Why/why not? (1p)

> For Q2b-f, you can write the solutions using formulas, words, Venn diagrams, code, numbers or the combination of them, whichever expresses your rationales the best. If you find handwriting more convenient, feel free to do so and attach a photo of it in the submission.

\vfill

*Have fun and good luck!*\
*Huong and Emelie*
