---
title: "NDH802 - Hypothesis testing"
author: "Chapter 9 and 10"
output: 
 prettydoc::html_pretty:
   theme: hpstr
   highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 9.55

*A random sample of 10 students contains the following observations, in hours, for time spent studying in the week before final exams:*

*28 57 42 35 61 39 55 46 49 38*

*Assume that the population distribution is normal.*

a.  Find the sample mean and standard deviation.

    ```{r}
    sample = c(28,57,42,35,61,39,55,46,49,38) #input the assumptions
    mu_0 = mean(sample) #we learnt this from R live session 1
    sd = sd(sample)  #we learnt this from R live session 1
    ```

b.  Test, at the 5% significance level, the null hypothesis that the population mean is 40 hours against the alternative that it is higher.

```{=tex}
\begin{align*}
H_o: \mu = 40\\
H_1: \mu > 40
\end{align*}
```
[R shortcuts:]{.ul}

```{r}
ttest = t.test(sample, mu = 40, alternative = "greater")
ttest$p.value
```

Because the p-value is higher than $\alpha =$ 0.05, we fail to reject the null hypothesis.

[By hand:]{.ul}

```{r}
n = 10
mu = 40

t = (mu_0 - mu)/(sd/sqrt(n)) #just plug in the numbers
t_score = qt(p = 0.95, df = n-1) #use the t-table as you prefer
```

Because `t_score` = `r t_score` is higher than `t` = `r t`, we fail to reject the null hypothesis.

## 9.37

The probability of type II error is:

$\beta = P(\bar{x}\le 5.041|\mu = \text{each of the mu given in the sub-questions})$

$\text{power} = 1 - \beta$.

```{r}
#we calculate this similarly to the way we solve exercises in previous chapters
beta_a = pnorm(q = 5.041, mean = 5.1, sd = 0.1/sqrt(16))
power_a = 1 - beta_a

beta_b = pnorm(q = 5.041, mean = 5.03, sd = 0.1/sqrt(16))
power_b = 1 - beta_b

beta_c = pnorm(q = 5.041, mean = 5.15, sd = 0.1/sqrt(16))
power_c = 1 - beta_c

beta_d = pnorm(q = 5.041, mean = 5.07, sd = 0.1/sqrt(16))
power_d = 1 - beta_d
```

## 9.65

```{=tex}
\begin{align*}
H_o: \mu \le 40 \\
H_1: \mu > 40
\end{align*}
```
```{r}
#summrize the assumption
n_965 = 125
xbar_965 = 40.9
s_squared_965 = 65

t_965 = (xbar_965 - 40)/sqrt(s_squared_965/n_965)
pvalue_965 = pt(q = t_965, df = n_965-1, lower.tail = FALSE)
```

p-value = `r pvalue_965` is not very small, hence the claim is not very strong.

## 10.4

```{=tex}
\begin{align*}
H_o: \mu_1 = \mu_2 \\
H_1: \mu_1 > \mu_2
\end{align*}
```
```{r}
#load data into your Global environment
price <- read.csv("https://raw.githubusercontent.com/lanhuongnguyen276/NDH802/master/Spring%202021/Exercises/HouseSellingPrice.csv")[, -1]



#t.test on all data
t.test(x = price$Sale.1.Price, 
       y = price$Sale.2.Price,
       #here we specify paired = TRUE because we want to do matched t-test
       paired = TRUE,
       #here we specify alternative = "greater" because that is our H1
       alternative = "greater")

#t.test on Atlanta data
t.test(x = price[price$Atlanta == 1, "Sale.1.Price"],
       y = price[price$Atlanta == 1, "Sale.2.Price"],
       paired = TRUE, alternative = "greater")
```

p-value in both cases are higher than the alpha we normally specify, therefore we fail to reject the null hypotheses.

## 10.34

Here I chose $\alpha = 0.05$ for example, feel free to choose other $\alpha$ as you reason.

```{r}
nutrition <- read.csv("https://raw.githubusercontent.com/lanhuongnguyen276/NDH802/master/Exercises/Food_Nutrition_Atlas.csv")[, -1]
#summary(nutrition)
```

```{=tex}
\begin{align*}
H_o: \mu_{adult-metro} = \mu_{adult-nonmetro}\\
H_1: \mu_{adult-metro} \ne \mu_{adult-nonmetro}
\end{align*}
```
```{r}
#The following code is simply data manipulation
#"PCT_OBESE_ADULTS" of adults living in metro cities
adult_metro = nutrition[nutrition$metro == 1, "PCT_OBESE_ADULTS"]
#"PCT_OBESE_ADULTS" of adults living in non-metro cities
adult_nonmetro = nutrition[nutrition$metro == 0, "PCT_OBESE_ADULTS"]
#perform the t test
t.test(x = adult_metro,
       y = adult_nonmetro,
       #we specify alternative = "two.sided" because H1 is "not equal"
       alternative = "two.sided")
```

p-value is smaller than alpha, we reject the null hypothesis.

```{=tex}
\begin{align*}
H_o: \mu_{child-metro} = \mu_{child-nonmetro}\\
H_1: \mu_{child-metro} \ne \mu_{child-nonmetro}
\end{align*}
```
```{r}
#This is similar, but on the children
child_metro = nutrition[nutrition$metro == 1, "PCT_Child_OBESITY"]
child_nonmetro = nutrition[nutrition$metro == 0, "PCT_Child_OBESITY"]
t.test(child_metro, child_nonmetro)
```

p-value is higher than alpha, we fail to reject the null hypothesis.

## 10.48

$\alpha = 0.01$

```{=tex}
\begin{align*}
H_o: \mu_{SalesO} = \mu_{SalesC}\\
H_1: \mu_{SalesO} \ge \mu_{SalesO}
\end{align*}
```
```{r}
ole <- read.csv("https://raw.githubusercontent.com/lanhuongnguyen276/NDH802/master/Exercises/Ole.csv")[, -1]

t.test(x = ole$Olesales,
       y = ole$Carlsale,
       alternative = "greater")
```

p-value \< alpha, we reject the null hypothesis.

```{=tex}
\begin{align*}
H_o: \mu_{PriceO} = \mu_{PriceC}\\
H_1: \mu_{PriceO} \ne \mu_{PriceO}
\end{align*}
```
```{r}
t.test(x = ole$Oleprice,
       y = ole$Carlpric,
       alternative = "two.sided")
```

p-value \> alpha, we fail to reject the null hypothesis.

## 10.52

Let $x_1, x_2$ denote the diets of the immigrants in the first and second interview; $y_1, y_2$ denote the diets of the non-immigrants in the first and second interview.

Here are the hypotheses we want to test for the first interview. You can try the second interview yourself.

```{=tex}
\begin{align*}
H_o: \mu_{x1} \le \mu_{y1}\\
H_1: \mu_{x1} > \mu_{y1}
\end{align*}
```
```{r}
#data prep
hei <- read.csv("https://raw.githubusercontent.com/lanhuongnguyen276/NDH802/master/Exercises/HEI.csv")[, -1]

x1 = hei[hei$immigrant == 1 & hei$daycode == 1, "HEI2005"]
y1 = hei[hei$immigrant == 0 & hei$daycode == 1, "HEI2005"]
x2 = hei[hei$immigrant == 1 & hei$daycode == 2, "HEI2005"]
y2 = hei[hei$immigrant == 0 & hei$daycode == 2, "HEI2005"]

#difference in diet from the first interview
t.test(x = x1, y = y1, alternative = "greater")

#difference in diet from the second interview
t.test(x = x2, y = y2, alternative = "greater")
```

## 10.56

Let $f_1, f_2$ denote the daily cost of women in the first and second interview; $m_1, m_2$ denote the daily cost of men in the first and second interview.

Here are the hypotheses we want to test for the first interview. You can try the second interview yourself.

```{=tex}
\begin{align*}
H_o: \mu_{f1} \ge \mu_{m1} \\
H_1: \mu_{f1} < \mu_{m1}
\end{align*}
```
```{r}
#data prep
f1 = hei[hei$female == 1 & hei$daycode == 1, "daily_cost"]
m1 = hei[hei$female == 0 & hei$daycode == 1, "daily_cost"]
f2 = hei[hei$female == 1 & hei$daycode == 2, "daily_cost"]
m2 = hei[hei$female == 0 & hei$daycode == 2, "daily_cost"]

#difference in daily cost from the first interview
t.test(x = f1, y = m1, alternative = "less")
#difference in daily cost from the seond interview
t.test(x = f2, y = m2, alternative = "less")
```
