---
title: "The old normal"
author: "NDH802 - Spring 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
if(!require(asbio)){
    install.packages("asbio")
    library(asbio)
}
```

### Exercise 5.27

First, type in the assumptions from the question. Here I divide the $\mu$ and $\sigma$ by 1000 for presentation purpose only. Therefore I will work on '000 SEK going forward, i.e., when you see 500, it means 500,000 SEK. You can totally skip this step if you want to.

```{r}
mu = 500000/1000
sigma = 50000/1000
```

**(a) What is the probability that the cost of building a home will be between \$460,000 and \$540,000? That is, find** $P(460,000 < X < 540,000)$ **where** $X$ **denotes the cost?**\
What we want to compute is the *blue* area. To do that in base R, one way is to compute the gray area, then subtract the pink area from it.

```{r, fig.height=7, fig.width=6, echo=FALSE, fig.align='center'}
{
  #This is for illustration purpose only. You don't have to learn to code this.
  par(mfrow = c(3, 1))
  shade.norm(
    from = mu - 4 * sigma,
    to = 540,
    sigma = sigma,
    mu = mu,
    tail = "middle",
    show.dist = F,
    show.p = F
  )
  shade.norm(
    from = mu - 4 * sigma,
    to = 460,
    sigma = sigma,
    mu = mu,
    tail = "middle",
    show.dist = F,
    show.p = F,
    shade.col = "pink3"
  )
  shade.norm(
    from = 460,
    to = 540,
    sigma = sigma,
    mu = mu,
    tail = "middle",
    show.dist = F,
    show.p = F,
    shade.col = "lightblue3"
  )
}
```

To compute the area below the normal curve, we use `pnorm()`.

```{r}
grey = pnorm(540, mean = mu, sd = sigma)
pink = pnorm(460, mean = mu, sd = sigma)
blue = grey - pink; cat(blue)
```

*Can you suggest another way to compute the blue area?*\
Hint: The total area below the normal curve is 1 and `pnorm()` can compute the area from either tails. Try `?pnorm` and check out the argument `lower.tail`.

```{r}
# blue_alternative <- 1 - #total area below the curve
#   pnorm(460, mean = mu, sd = sigma) - #the white area on the left
#   pnorm(540, mean = mu, sd = sigma, lower.tail = FALSE) #the white area on the right
# cat(blue_alternative)
```

**(b) What probability is 0.2 that the cost of building will be less than what amount? That is, find** $a$ **such that** $P(X < a) = 0.2$**?**

```{r}
a = qnorm(0.2, mean = mu, sd = sigma)
cat(a)
```

**(c) Find the shortest range such that the probability is 0.95 that the cost of a luxury home will fall in this range. That is, find** $b$ **and** $c$ **such that** $P(b < X < c) = 0.95$**?**\
The graph is, again, for illustration purpose, you don't need to learn to code it. Notice the lavender area is 0.95.

```{r, fig.height=3, fig.width=6, fig.align='center'}
b = qnorm(0.025, mean = mu, sd = sigma)
c = qnorm(0.975, mean = mu, sd = sigma)
```

```{r, echo = FALSE, fig.height=3, fig.width=6, fig.align='center'}
shade.norm(
  from = b,
  to = c,
  sigma = sigma,
  mu = mu,
  tail = "middle",
  show.dist = F,
  show.p = F,
  shade.col = "lavender"
)
```

***NOTE***\
- In (a), we compute the **area** below the curve (the gray/pink/blue area) which characterizes the **probability**.\
- In (b) and (c), we compute the **value** of $X$ (a point on the x-axis) which characterizes **the cost of building a home**.
