---
title: "NDH802 R Application - Live session 2"
author: "Huong Nguyen"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Live session

```{r}
salaries = read.csv("https://bit.ly/3r918BW")
```

## Probability of an event

From the data you have, what is the probability of RM/BE student? More formally, compute $P(RM)$ and $P(BE)$.

```{r}
table(salaries$program)/nrow(salaries)
prop.table(table(salaries$program))
#these two commands give the same results
```

Similarly, can you compute $P(male)$ and $P(female)$?

```{r}
table(salaries$whatisyourgender1isfemale)/nrow(salaries) #including NA
```

```{r}
prop.table(table(salaries$whatisyourgender1isfemale)) #without NA observations
```

## Joint and conditional probability

Now let's spice things up a little bit. Assume we want to compute $P(RM \cap female)$, $P(female\mid RM)$ and $P(RM \mid female)$

```{r}
#If we don't put anything, the sample space is all of the observations included in our table
a = prop.table(table(salaries$program, salaries$whatisyourgender1isfemale))
sum(a) #all of them add up to 1
```

As some of you asked, you can write like this in the assignment.

$P(RM \cap female)$ is

```{r, echo=FALSE}
print(a[1,1]) 
```

```{r}
#If you put margin =1, we change the sample space to the variable in the rows (in this case, the program BE/RM)
b = prop.table(table(salaries$program, salaries$whatisyourgender1isfemale),
           margin = 1)
sum(b[1,]) #the probabilities in each row add up to 1
sum(b[2,]) #the probabilities in each row add up to 1
```

```{r}
#If you put margin =2, we change the sample space to the variable in the columns (in this case, the gender female/male)
c = prop.table(table(salaries$program, salaries$whatisyourgender1isfemale),
           margin = 2)
sum(c[,1]) #the probabilities in each column add up to 1
sum(c[,2]) #the probabilities in each column add up to 1
```

# Your turn:

```{r}
survey = read.csv("https://raw.githubusercontent.com/lanhuongnguyen276/NDH802/master/Live%20sessions/NDH802survey.csv")[,-1]
```

## Probability of an event

$P(Q2 = \text{R})$

```{r}
#We learn this in the morning session
prop.table(table(survey$Q2))

#If you want to "extract" a number from your prop.table
R = prop.table(table(survey$Q2))[2] #because R is the second element
```

$P(Q1 = \text{Love it})$

```{r}
#We learn this in the morning session
prop.table(table(survey$Q1))

#If you want to "extract" a number from your prop.table
loveit = prop.table(table(survey$Q1))[2] #because loveit is the second element
```

$P(Q1 = \text{Crying} \cup \text{Meh})$

Formally, we learn that $P(Crying \cup Meh) = P(Crying) + P(Meh) - P(Crying \cap Meh)$. We know that Crying and Meh are mutually exclusive i.e. $P(Crying \cap Meh) = 0$. Therefore, $P(Crying \cup Meh) = P(Crying) + P(Meh)$

```{r}
crying = prop.table(table(survey$Q1))[1]
meh = prop.table(table(survey$Q1))[3]
crying_union_meh = as.numeric(crying) + as.numeric(meh) #as crying and meh are not numbers, we "force" them to be numbers with as.numeric
```

We also know that Q1 has only 3 outcomes (Love it, Meh and Crying) therefore they are collectively exhaustive (and by construction, they are mutually exclusive). Therefore, $P(Crying) + P(Love it) + P(Meh) = 1$

```{r}
#Another way to calculate crying_union_meh
crying_union_meh_2 = 1 - as.numeric(loveit) #notice that it gives you the same results as before
```

## Conditional probability

$P(Q2 = \text{R}\mid Q1 = \text{Love it})$

```{r}
#This code gives you the prop.table
prop.table(table(survey$Q1, survey$Q2), margin = 1)

#Then you can extract speciic values from your prop.table
R_given_loveit = prop.table(table(survey$Q1, survey$Q2), margin = 1)[2,2] #because R|love it is second row/second colum
```

$P(Q2 = \text{Excel}\mid Q1 = \text{Love it})$

```{r}
Excel_given_loveit = prop.table(table(survey$Q1, survey$Q2), margin = 1)[2,1] #because R given love it is second row/second colum
```

Notice that $P(Q2 = \text{R}\mid Q1 = \text{Love it}) + P(Q2 = \text{Excel}\mid Q1 = \text{Love it}) = 1$

```{r}
R_given_loveit + Excel_given_loveit
```

### Joint probability

$P(Q2 = \text{R}\cap Q1 = \text{Love it})$

```{r}
#This code gives you the prop.table
prop.table(table(survey$Q1, survey$Q2))

#Then you can extract speciic values from your prop.table
R_and_loveit = prop.table(table(survey$Q1, survey$Q2))[2,2] #because R and love it is second row/second colum
```

$P(Q2 = \text{Excel}\cap Q1 = \text{Love it})$

```{r}
Excel_and_loveit = prop.table(table(survey$Q1, survey$Q2))[2,1] #because Excel and love it is second row/first colum
```

Notice that $P(Q2 = \text{Excel}\cap Q1 = \text{Love it}) + P(Q2 = \text{R}\cap Q1 = \text{Love it}) = P(Q1 = \text{Love it})$

```{r}
print(paste(loveit, R_and_loveit + Excel_and_loveit))
#you don't need to learn to code print() and paste(). It is for illustration purpose only
```

$P(Q2 = \text{R}\cap Q1 = \text{Crying} \cup \text{Meh})$

```{r}
as.numeric(R) - R_and_loveit
#I'll let you think about why :) But please don't do it to me in the assignments. Do explain why, it's the most exciting part.
```
