---
title: "NDH802 R Application - Live session 2"
author: "Huong Nguyen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Set your working directory

```{r}
setwd("~/Documents/GitHub/NDH802") #This is my WD, change it to yours!
```

### Load your data

```{r}
salaries = read.csv("https://bit.ly/3r918BW")
```

### Discrete distribution

From the data you have, what is the probability of RM/BE student? More formally, compute $P(RM)$ and $P(BE)$.

```{r}
table(salaries$program)
prop.table(table(salaries$program))
```

Similarly, can you compute $P(male)$ and $P(female)$?

```{r}

```

Now let's spice things up a little bit. Assume we want to compute $P(RM \cap female)$, $P(female\mid RM)$ and $P(RM \mid female)$

```{r}
#margin = 1 means by row, margin = 2 means by column
prop.table(table(salaries$program, salaries$whatisyourgender1isfemale))
prop.table(table(salaries$program, salaries$whatisyourgender1isfemale),
           margin = 1)
prop.table(table(salaries$program, salaries$whatisyourgender1isfemale),
           margin = 2)
```

### Your turn:

-   Reserve a slot for the Q&A

-   Go through Assignment 2 and try what you can
