---
title: "NDH802 - Hypothesis testing"
author: "Huong"
output:
  prettydoc::html_pretty:
    theme: hpstr
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 9.55

*A random sample of 10 students contains the following observations, in hours, for time spent studying in the week before final exams:*

*28 57 42 35 61 39 55 46 49 38*

*Assume that the population distribution is normal.*

a.  Find the sample mean and standard deviation.

    ```{r}
    sample = c(28,57,42,35,61,39,55,46,49,38) #input the assumptions
    mu_0 = mean(sample) #we learnt this from R live session 1
    sd = sd(sample)  #we learnt this from R live session 1
    ```

b.  Test, at the 5% significance level, the null hypothesis that the population mean is 40 hours against the alternative that it is higher.

```{=tex}
\begin{align*}
H_o: \mu =40\\
H_1: \mu>40
\end{align*}
```
[R shortcuts:]{.ul}

```{r}
ttest = t.test(sample, mu = 40, alternative = "greater")
ttest$p.value
```

Because the p-value is higher than 0.05, we fail to reject the null hypothesis.

[By hand:]{.ul}

```{r}
n = 10
mu = 40

t = (mu_0 - mu)/(sd/sqrt(n)) #just plug in the numbers
t_score = qt(p = 0.95, df = n-1) #use the t-table as you prefer
```

Because `t_score` is higher than `t`, we fail to reject the null hypothesis.
