---
title: "NDH802 R Application - Live session 1"
author: "Huong Nguyen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load your data

```{r}
salaries = read.csv("https://bit.ly/3r918BW") #data stored on cloud
```

## Box plot part

```{r, echo=FALSE}
b = boxplot(salaries[,"sal_expected"]/1000, ylim = c(0, 100)) #a plot like this in the assignment will get 0.5pt
```

The below code give you the statistics of the box plot, in order

1.  **the "minimum"**: Q1 -1.5\*IQR

2.  **the first quartile (Q1/25th Percentile)**: the middle number between the smallest number (not the "minimum") and the median of the dataset.

3.  **the median (Q2/50th Percentile)**: the middle value of the dataset.

4.  **the third quartile (Q3/75th Percentile)**: the middle value between the median and the highest value (not the "maximum") of the dataset.

5.  **the "maximum"**: Q3 + 1.5\*IQR

    where **interquartile range (IQR)** is 25th to the 75th percentile.

```{r}
b$stats
```
