---
title: "NDH802 R Application - Seminar #2"
author: "Huong Nguyen"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data in

```{r}
salaries = read.csv("https://bit.ly/3r918BW")
```

Let $A$ be an event. There are several ways to calculate $N_A$ in base R, today we learn 3 ways. For example, assume we want to know the number of BE ($N_{BE}$) and RM ($N_{RM}$) students in our data, we can

1.  Count the number of rows that contain BE and RM students.

    ```{r}
    N_RM = nrow(salaries[salaries$program == "RM",])
    N_BE = nrow(salaries[salaries$program == "BE",])
    ```

2.  Use logic, that is see if the program is RM or BE.

    ```{r}
    N_RM = sum(salaries$program == "RM")
    N_BE = sum(salaries$program == "BE")
    ```

3.  To compute several events at the same time, `table()` can be handy.

    ```{r}
    table(salaries[,"program"])
    ```

    We can do something similar to a pivot table with `table()`

    ```{r}
    table(salaries[,"program"], salaries[,"whatisyourgender1isfemale"])
    ```

    However, it could be a bit messy when the number of events is large. For example:

    ```{r}
    table(salaries[,"worktime"])
    ```

### Your turn:

(1) How many male students and female students do we have in the data set? Formally, compute $N_{male}$ and $N_{female}$

(2) How many *students who are female and belong to RM program* do we have in the data set? compute $N_{female \cap RM}$

## Probability of an event

Recall from the lecture,

$$
P(A) = \frac{N_A}{N}
$$

From the data you have, what is the probability of RM/BE student? More formally, compute $P(RM)$ and $P(BE)$.

$$
P(RM) = \frac{N_{RM}}{N}; P(BE) = \frac{N_{BE}}{N}
$$

As we are going to use $N$ repeatedly, it's better to ask R to remember it.

```{r}
N <- nrow(salaries)
```

Now, to calculate $P(RM)$, we also need $N_{RM}$, which we already had. Then

```{r}
P_RM = N_RM/N
```

Similarly,

```{r}
P_BE = N_BE/N
```

We can also do like this:

```{r}
table(salaries$program)/nrow(salaries)
prop.table(table(salaries$program))
#these two commands give the same results
```

### Your turn:

Can you compute $P(male)$ and $P(female)$?

## Joint and conditional probability

Now let's spice things up a little bit. Assume we want to compute $P(RM \cap female)$, $P(female\mid RM)$ and $P(RM \mid female)$. Similarly,

$$
P(RM \cap female) = \frac{N_{RM \cap female}}{N}
$$

By this formula, we first need to calculate $N_{RM \cap female}$, using any ways discussed earlier.

```{r}

```

$$
P(RM \mid female) = \frac{P(RM \cap female)}{P(female)} = \frac{N_{RM \cap female}/N}{N_{female}/N} = \frac{N_{RM \cap female}}{N_{female}} 
$$

```{r}

```

### Your turn:

1.  Compute $P(RM \mid female)$
2.  Compute $P(RM \cup female)$
3.  Why is $P(RM \mid female)$ different from $P(female \mid RM)$?
