---
title: 'NDH802 R Application - Seminar #1'
author: "Huong Nguyen"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Working with data in R

### Load your data

If you are familiar with Excel, this step is like opening a file. You can type in

```{r}
salaries = read.csv("https://tinyurl.com/NHD802salaries") #data stored on cloud
```

or click on Import Data set on the top right of your screen.

### Explore your data

These commands give you a quick overview of your data. I recommend you to try and run them in Rstudio, but NOT to knit them. It's gonna be a bit messy. To mask a command, we make them a "comment" by placing \# in front of the command. Things after \# are not code and won't be executed.

```{r, results='hide'}
#View(salaries) 
head(salaries)
tail(salaries)
summary(salaries)
#rename a column
```

### Select columns from a data frame

There are numerous way to select a column. The three most basic way are:

```{r, results='hide'}
#here you refer to the column by the columns names
salaries$sal_expected
salaries[,"sal_expected"]

#here you refer to the column by the column position from the left ie 1 is the first column from the left
salaries[,2] 
```

**When to use which?**\
It's all about your personal preference. I introduce several ways to give you options, **not** to confuse you. You can totally explore, see which you're more comfortable with and stick with it. This is also applicable for most of other commands.

To select more than 1 column:

```{r, results='hide'}
#here you refer to the column by the columns names
salaries[,c("sal_expected", "fairpay", "program")]

#here you refer to the column by the column position from the left
salaries[,c(1,3,10)]  
```

### Select rows from a data frame

There are numerous way to select a row. The most basic way (with base R) are:

```{r, results='hide'}
salaries[salaries$program == "RM",]
salaries[salaries$whatisyourgender1isfemale == 1,]
```

### Let's take a closer look

Assume you want to explore the salary expectation from the students, which is the variable `sal_expected`. Two of my most frequent used methods are histogram and box plot.

Let's first try plotting a **histogram**. If you're like "a histo what?", take a quick look [here](https://www.youtube.com/watch?v=qBigTkBLU6g), then come back.

```{r}
hist(salaries[,"sal_expected"]) #this in the assignment will get 0.5pt
```

This looks neither aesthetically appealing nor informative, right? How do you think we can make it better?

```{r, echo=FALSE, results='hide'}
hist( #I expect something like this
  salaries[,"sal_expected"] / 1000,
  xlim = c(0, 100),
  breaks = 200,
  xlab = "expected salary ('000 SEK)",
  main = "Expected Salary",
  col = "lightpink3"
)
```

Now let's try **box plot**. Similarly, if you need a quick understanding of box plot, take a look [here](https://www.youtube.com/watch?v=7UK2DK7rblw), then come back.

```{r, echo=FALSE}
boxplot(salaries[,"sal_expected"]) #a plot like this in the assignment will get 0.5pt
```

This looks nothing like the video. Why do you think it is and can you help me fix it?

```{r, echo=FALSE, results='hide'}
boxplot( #I expect something like this
  sal_expected ~ program,
  data = salaries,
  ylim = c(0, 100000)
)
```

## Means and standard deviation (SD)

### Calculate the mean and SD of a variable

Intuitively, you would type

```{r}
mean(salaries[,"sal_expected"])
sd(salaries[,"sal_expected"])
```

but it doesn't work. why? Because we have NA in our data. To ask R to ignore the NA values,

```{r}
mean(salaries[,"sal_expected"], na.rm = TRUE)
sd(salaries[,"sal_expected"], na.rm = TRUE)
```

### Can I calculate the mean of salary expected of the RM, and the mean of salary expected of the BE students separately?

```{r}
mean(salaries[salaries$program == "BE","sal_expected"], na.rm = TRUE)
mean(salaries[salaries$program == "RM","sal_expected"], na.rm = TRUE)
```

### Can I calculate the mean of salary expected of the female RM student?

```{r}
mean(salaries[salaries$program == "BE" & salaries$whatisyourgender1isfemale == "1", "sal_expected"], na.rm = TRUE)
```

## Your turn

(1) Plot a histogram for `worktime`.

    ```{r}
    hist(salaries$worktime, main = "Work time", xlab = "Work time")
    ```

(2) Make a box plot for `idealpay`.

    ```{r}
    boxplot(salaries$idealpay/1000, ylim = c(0, 100), ylab = "'000 SEK", main = "Ideal Pay")
    ```

(3) Make a box plot for `idealpay`, for RM and BE students separately.

    ```{r}
    boxplot(idealpay/1000 ~ program, data = salaries, ylim = c(0,100),
            main = "Ideal Pay",
            ylab = "'000 SEK")
    ```

(4) Compute the mean and variance of `fairpay`.

    ```{r}
    mean(salaries$fairpay, na.rm = TRUE)
    var(salaries$fairpay, na.rm = TRUE)
    ```

(5) Compute the mean and variance of `fairpay` of the female students.

    ```{r}
    mean(salaries[salaries$whatisyourgender1isfemale == 1, "fairpay"], na.rm = TRUE)
    var(salaries[salaries$whatisyourgender1isfemale == 1, "fairpay"], na.rm = TRUE)
    ```

(6) Compute the mean and variance of `fairpay` of the students who are female and belong to RM program.

    ```{r}
    mean(salaries[salaries$program == "RM" & salaries$whatisyourgender1isfemale == 1, "fairpay"], na.rm = TRUE)
    var(salaries[salaries$program == "RM" & salaries$whatisyourgender1isfemale == 1, "fairpay"], na.rm = TRUE)
    ```

## Or,

try Assignment 1, Question 1.
